### 2017-04-04 glossaries.tex
Add glossary

\usepackage{datatool}
\usepackage[acronym,toc]{glossaries}
\makeglossaries

\newacronym{hab}{HAB}{High-Altitude Balloon}
\newacronym{gnss}{GNSS}{Global Navigation Satellite System}
\newacronym{obc}{OBC}{On Board Computer}
\newacronym{ahabus}{AHABus}{Abertay High-Altitude Bus}
\newacronym{fec}{FEC}{Forward Error Correction}
\newacronym{i2c}{I\textsuperscript{2}C}{Inter-Integrated Circuit}
\newacronym{ccsds}{CCSDS}{Consultative Committee for Space Data Systems}
\newacronym{ecc}{ECC}{Error-Correction Codes}
\newacronym{rtty}{RTTY}{Radio-Teletype}