### 2017-04-07 paper.tex

\input{header.tex}
\input{glossaries.tex}

\begin{document}

\frontmatter

\begin{titlepage}
\maketitle
\end{titlepage}

\begin{abstract}
\end{abstract}

\tableofcontents
\setlength{\parskip}{1em}

\listoffigures

\listoftables

\printglossary
\printglossary[type=\acronymtype]

\mainmatter
\input{chapters/01-introduction.tex}

\newpage
\input{chapters/02-lit-review.tex}

\newpage
\input{chapters/03-methodology.tex}

\newpage
\input{chapters/04-results.tex}

\newpage
\input{chapters/05-discussion.tex}

\newpage
\input{chapters/06-conclusion.tex}

\printbibliography[heading=bibintoc, title=References]

\end{document}