### 2017-04-08 chapters/05-discussion.tex
Add Aknowledgements

\chapter{Discussion}
\label{ch:discussion}

\section{Introduction}

The aim of the project was to design and build a prototype of a High-Altitude
Balloon modular payload bus, similar in principle to buses used in the
satellite industry, and evaluate both the feasibility of such a platform and
its performance.

The AHABus platform prototype was

\section{Project Evaluation}

\subsection{Overview}

\subsection{Radio Communications System}

\subsection{Payload Bus}

\subsection{Ground Segment}

\section{Analysis of Test Results}

\subsection{Day-in-the-Life Test}

\subsection{Radio System Resilience Tests}

\section{Further Work} % << mix in with the rest?