### 2017-04-08 paper.tex

\documentclass[11pt]{report}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{titlesec}
\usepackage{float}
\usepackage{color}
\usepackage{listings}
\usepackage{lmodern}
\usepackage{graphicx}
\usepackage{bytefield}
\usepackage{pgfplots}
\usepackage{csvsimple}
\usepackage[detect-weight=true, binary-units=true]{siunitx}
\usepackage{tikz}
\usepackage[justification=centering,labelfont={small,bf},font={small,bf}]{caption}
\usepackage[scaled]{beramono} % BitStream Vera font (code)
\usepackage[T1]{fontenc}
\usepackage[a4paper, margin=2.54cm]{geometry}
\usepackage[nottoc]{tocbibind}

%\renewcommand{\familydefault}{\sfdefault}
\pgfplotsset{width=10cm,height=6cm,compat=newest}
\graphicspath{{figures/}}

\usepackage[
style=authoryear,
citestyle=authoryear,
backend=biber,
firstinits=true
]{biblatex}
\DeclareNameAlias{author}{last-first}

\addbibresource{../library.bib}
\DefineBibliographyStrings{english}{%
  references = {References},
}

\lstset{basicstyle=\footnotesize\ttfamily,breaklines=true}

\author{Amy Parent}
\title{AHABus \& FCORE:\\Modular Bus for High-Altitude Balloon Missions}
\date{April 2017}

% Parameters

\setlength{\baselineskip}{1.5em}

% Title

\newcommand\frontmatter{%
    \cleardoublepage
    \pagenumbering{roman}}

\newcommand\mainmatter{%
    \cleardoublepage
    \pagenumbering{arabic}}

\makeatletter
\def\maketitle{%
    \vspace*{10em}
	\begin{center}%
	\let \footnote \thanks
        \vskip 5em%
		{\LARGE \textbf{\@title} \par}%
		\vskip 1.5em%
		{\large \@author}%
		\vskip 1em%
        {\large BSc (Hons) Computing - \@date\\} %
		\vskip 1em%
		{\normalsize School of Arts, Media and Games\\}%
		\vskip .1em%
		{\normalsize University of Abertay Dundee\\}%
		\vskip .1em%
		{\normalsize Dundee, DD1 1HG, UK}%
	\end{center}
}
\makeatother

\input{glossaries.tex}

\begin{document}

\frontmatter

\begin{titlepage}
\maketitle
\end{titlepage}

\begin{abstract}
\end{abstract}

\tableofcontents
\setlength{\parskip}{1em}

\listoffigures

\listoftables

\printglossary
\printglossary[type=\acronymtype]

\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

I would like to thank my supervisor, Dr. Ian Ferguson, for the support and
insight he gave throughout this research project, and as Dr. Natalie Coull for
her advice during the proposal phase of the project.

\mainmatter
\input{chapters/01-introduction.tex}

\newpage
\input{chapters/02-lit-review.tex}

\newpage
\input{chapters/03-methodology.tex}

\newpage
\input{chapters/04-results.tex}

\newpage
\input{chapters/05-discussion.tex}

\newpage
\input{chapters/06-conclusion.tex}

\printbibliography[heading=bibintoc, title=References]

\end{document}