### 2017-04-18 appendices.tex
First draft, minus photos

\begin{appendices}
%\addtocontents{toc}{\protect\setcounter{tocdepth}{0}}

\chapter{Source Code Repositories}
\label{apx:source-repos}

All source code and documentation for the \acrfull{ahabus} platform and, its flight software \acrfull{fcore} and the ground segment software are made available and grouped under the AHABus GitHub organisation, at https://github.com/ahabus. Below is a list of the different repositories:

\begin{description}
\item[\acrshort{fcore}] main flight software, based on FreeRTOS.\\https://github.com/ahabus/fcore

\item[Dl-Fldigi] Fork of Dl-Fldigi able to export raw RTTY binary streams.\\https://github.com/ahabus/dl-fldigi

\item[PacketRadio] AHABus Radio Protocol encoder/decoder reference implementation.\\https://github.com/AHABus/packet-radio
\end{description}

\chapter{Test Data Repositories}
\label{apx:test-data-repos}

The data collected during the testing phase is made available under the AHABus GitHub organisation. Below is a list of the different repositories:

\begin{description}
\item[Raw Data] https://github.com/ahabus/test-rawdata

\item[Data Processing Scripts] https://github.com/ahabus/test-processing
\end{description}

\end{appendices}