### 2017-04-18 paper.tex

\documentclass[11pt]{report}
\usepackage[scaled]{beramono} % BitStream Vera font (code)
\usepackage[T1]{fontenc}
\usepackage[a4paper, margin=3cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{titlesec}
\usepackage{float}
\usepackage{color}
\usepackage{listings}
\usepackage{lmodern}
\usepackage{graphicx}
\usepackage{bytefield}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{csvsimple}
\usepackage{hyperref}

\usepackage{mathptmx}

\usepackage[detect-weight=true, binary-units=true]{siunitx}
\usepackage{tikz}
\usepackage[justification=centering,labelfont={small,bf},font={small,bf}]{caption}
\usepackage[titletoc,toc,page]{appendix}
\usepackage[nottoc]{tocbibind}
\usepackage{array}

%\renewcommand{\familydefault}{\sfdefault}
\pgfplotsset{width=10cm,height=6cm,compat=newest}
\graphicspath{{figures/}}

\usepackage[
style=authoryear-comp,
citestyle=authoryear-comp,
backend=biber,
firstinits=true
]{biblatex}
\DeclareNameAlias{author}{last-first}

\addbibresource{../library.bib}
\DefineBibliographyStrings{english}{%
  references = {References},
}

\lstset{basicstyle=\footnotesize\ttfamily,breaklines=true}

\author{Amy Parent}
\title{AHABus \& FCORE: A Modular Bus System for High-Altitude Balloon Missions}
\date{April 2017}

% Parameters

\setlength{\baselineskip}{1.5em}
\setcounter{tocdepth}{4}
\setcounter{secnumdepth}{4}

% Title

\newcommand\frontmatter{%
\cleardoublepage
\pagenumbering{roman}
}

\newcommand\mainmatter{%
\cleardoublepage
\pagenumbering{arabic}
}

\makeatletter
\def\maketitle{%
    \vspace*{10em}
	\begin{center}%
	\let \footnote \thanks
        \vskip 5em%
		{\LARGE \textbf{\@title} \par}%
		\vskip 1.5em%
		{\large \@author}%
		\vskip 1em%
        {\large BSc (Hons) Computing - \@date\\} %
		\vskip 1em%
		{\normalsize School of Arts, Media and Games\\}%
		\vskip .1em%
		{\normalsize University of Abertay Dundee\\}%
		\vskip .1em%
		{\normalsize Dundee, DD1 1HG, UK}%
	\end{center}
}
\makeatother

\input{glossaries.tex}

\begin{document}

\frontmatter

\begin{titlepage}
\maketitle
\end{titlepage}

\begin{abstract}

The aim of this project was to investigate the design options for a modular payload bus for \acrfull{hab} missions, and to evaluate one such design by implementing in in hardware and software.

\acrlong{hab} provide affordable access to near-space conditions, even for low-budget entities (universities, schools and amateurs). While current low-budget missions rely on commercial off-the-shelf computers and sensors, their architecture and software are often custom-made and targeted for a single use. This prevents easy reuse of hardware and software components, and raises the barrier to entry by requiring non-trivial network and computing skills.

This paper presents the development of \acrfull{ahabus}, solution designed by applying the modular bus model used by most modern satellites to \acrlong{hab} architectures. A common, generic \acrfull{obc} running a common flight software (\acrshort{fcore}) provides facilities for position tracking and reliable radio telemetry to a ground station, and polls data from payloads, custom for each flight.

The resulting protocols and platform prototype were successful and passed both day-in-the-life testing and radio channel resilience tests. Further testing of the and minor modifications to the Payload Bus and Radio Communications protocols will be required in order to qualify the platform for in-flight use. Once these steps are taken, \acrshort{ahabus} could render scientific \acrlong{hab} missions more accessible and affordable by allowing students and researchers to focus on payload design and development.

\end{abstract}

\tableofcontents
\setlength{\parskip}{1em}

\listoffigures

\listoftables

\printglossary
\printglossary[type=\acronymtype]

\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

I would like to thank my supervisor, Dr. Ian Ferguson, for the support and
insight he gave throughout this research project, and as Dr. Natalie Coull for
her advice during the proposal phase of the project.

\mainmatter
\input{chapters/01-introduction.tex}

\newpage
\input{chapters/02-lit-review.tex}

\newpage
\input{chapters/03-methodology.tex}

\newpage
\input{chapters/04-results.tex}

\newpage
\input{chapters/05-discussion.tex}

\newpage
\input{chapters/06-conclusion.tex}

\printbibliography[heading=bibintoc, title=References]

\input{appendices.tex}

\end{document}